---
title: "STA130 R Simple Linear Regression"
author: "Scott Schwartz"
date: "Fall 2022 Week 8"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, include=TRUE, echo=TRUE)
```

# Part I: Data

```{r}
library(tidyverse)
#install.packages("Stat2Data")
library(Stat2Data) # This package contains the NCbirths data
# Sometimes, we need to use the data() function to load data from an R package into our R session
data(NCbirths) # After running this line, you'll see it appear in the Environment pane in the top left :)
glimpse(NCbirths)
```

## Q1 - What does `?NCbirths` show? 

###(a) How many observations are there in these data?

###(b) What does each observation represent?

###(c) Who contributed these data?

## Q2 - Add a `BirthWeightLbs` variable to the data using `BirthWeightOz` and the fact that there are 16 oz in 1 lb.

```{r}

```

## Q3 - Create a visualization to explore the association between the baby's weight at birth (on the y-axis) and gestational length (duration of a pregnancy in weeks) (on the x-axis). Describe this distribution.

```{r}

```

## Q4 - Consider the association between the duration of pregnancy and birthweight for babies born at 34 weeks or later. Create a new tibble containing only observations for babies born in this range, and plot the data association between `BirthWeightLbs` and `Weeks` now.

```{r}
# Remove babies born before 34 seeks who are quite premature
NCbirths_34plus <- NCbirths %>% 

glimpse(NCbirths_34plus)

# plot the data association on the restricted data set
```

# Part II: Simple Linear Regression

## Q1 - What's the correlation between `BirthWeightLbs` and `Weeks` for the restricted `NCbirths_34plus` data set? 
```{r}
# Does the correlation reflect our intuition about the association between these two variables?
```

## Q2 Create a visualization to visualize of association between weeks and birthweight for babies born at or after 34 weeks, adding the fitted regression line to the plot.
```{r}

```

## Q3 - Use the `lm` function to fit a linear regression model to find the least-squares parameter estimates of a simple linear regression predicting the weight of a baby based on the duration of the pregnancy (in weeks). 
```{r}

```

## Q4 - Interpret the estimated slope $\hat \beta_1$

## Q5 - Interpret the estimated intercept $\hat \beta_0$.

## Q6 - Based on the `NCbirths_34plus` data, is there evidence that the slope of this linear regression model significantly different from 0?

## Q7 - Calculate the proportion of the variability in weight that is explained by the simple linear regression model in two ways. 

```{r}
# What does this suggest?

```

# Part III: Indicator Variables

## Q1 - Create a visualization to explore the association between mothers' smoking status and birthweight (after 34 weeks), and describe what we observe.

```{r}

## What's weird here?
## 1. We get a warning saying the x aesthetic is continuous (when it should be categorical instead)
## 2. Some observations have missing values so are omitted

NCbirths_34plus <- NCbirths_34plus %>% filter(!is.na(Smoke) & !is.na(BirthWeightLbs)) %>%
  mutate(Smoke = case_when(Smoke == 0 ~ "Non-smoking mother",
                           Smoke == 1 ~ "Smoking mother"))
```

## Q3 - Fit a linear regression model to predict the birthweight of a baby based on his/her mother's smoking status.

```{r}

```

## Q3 - What is the baseline group in the model above?

## Q4 - Interpret the estimated intercept $\hat \beta_0$.

```{r}
# Checking our work
NCbirths_34plus %>% group_by(Smoke) %>%
  summarise(mean = mean(BirthWeightLbs))
# The numbers here match up with the values we get from the fitted regression model :)
```

## Q5 - Interpret the estimated "slope" $\hat \beta_1$ (which is actually a "contrast" rather than a "slope").

## Q6 - Based on these data, is there a difference between the mean weight of babies born to smoking vs non-smoking mothers?

# Part IV: Conclusion

Moving forward we'll  build richer linear regression models with more than 1 predictor. For example, it's clear based on these examples that there are many factors that affect an infants babies' birthweight, and looking at these potential predictors one at a time is not enough to make good predictions.
